---
title: "Lasso estimation in multiple linear regression"
author: "Gregoire Gasparini, Aurora Hofman, Sarah Musiol, Beatriu Tort"
date: "03 de marzo de 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Lasso for the Boston Housing data 
The Boston House-price dataset concerns housing values in 506 suburbs of Boston corresponding to year 1978. They are available here:
   https://archive.ics.uci.edu/ml/datasets/Housing

The Boston House-price corrected dataset (available in boston.Rdata) con- tains the same data (with some corrections) and it also includes the UTM coor- dinates of the geographical centers of each neighborhood.

### Lasso estimation using the package 'glmnet'
After loading the right package, the response and explanatory variables from the Boston Housing data are set.
```{r}
#install.packages("glmnet")
library(glmnet)

boston <- load("boston.Rdata")

response <- "CMEDV"

explanatory <- c("CRIM", "ZN", "INDUS", "CHAS", "NOX", "RM", "AGE", "DIS", "RAD", "TAX", "PTRATIO", "B", "LSTAT")

# glmnet cannot handle factors -> "CHAS" is a factor
boston.c$CHAS <- as.numeric(boston.c$CHAS)
```


### 



## 2. A regression model with $p >> n$

```{r}

```

