---
title: "Local_poison_regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bandwidth choice for the local Poisson regression

The functions from ATENEA are modified to fit a local Poisson regression.

```{r}

#The first function is not modified
h.cv.sm.poison <- function(x,y,rg.h=NULL,l.h=10,method=loglik.CV){
   cv.h <- numeric(l.h)
   if (is.null(rg.h)){
      hh <- c(h.select(x,y,method="cv"),
            h.select(x,y,method="aicc"))
      rg.h <- range(hh)*c(1/1.1, 1.5)
   }
   i <- 0
   gr.h <- exp( seq(log(rg.h[1]), log(rg.h[2]), l=l.h))
   for (h in gr.h){
      i <- i+1
      cv.h[i] <- method(x,y,h)
   }
   return(list(h = gr.h, 
               cv.h = cv.h, 
               h.cv = gr.h[which.min(cv.h)]))
}


# method loglik.CV: leave-one-out log-likelihood 

#Modifications: sm.poisson and the return function is the loglikelihood of the density of the poison distribution. 

loglik.CV <- function(x,y,h){
  n <- length(x)
  pred <- sapply(1:n, 
      function(i,x,y,h){
         sm.poisson(x=x[-i],y=y[-i],h=h,eval.points=x[i],display="none")$estimate
      },   x,y,h)
  #faculty <- sapply(1:n, function(i, y){prod(1:y[i])}, # if one wants the scalingfactor uncomment this line
  return(-sum(y*log(pred)-pred)) #-log(faculty) only a scale 
  
}

```

# Local Poisson regression for Country Development Data

Consider the country development dataset containing information on development indicators measured in 179 countries.

```{r}
countries<-read.csv2(file="HDI.2017.subset.csv",row.names = 1)
#attach(countries)
le.fm.r <- round(countries$le.fm)
```

Fit a local Poisson regression modeling le.fm.0 as a function of Life.expec. 
Use sm.poisson from R package.

```{r}
library(sm)

x<- countries$Life.expec
y<- le.fm.r

h.cv.sm.poison(countries$Life.expec, le.fm.r)$h.cv

```

Plot of the cross-validation as a function of Bandwidth.

```{r}
h.CV.loglik <- h.cv.sm.poison(countries$Life.expec,le.fm.r, l.h = 10)
plot(h.CV.loglik$h,h.CV.loglik$cv.h)
lines(h.CV.loglik$h,h.CV.loglik$cv.h)
abline(v=h.CV.loglik$h.cv)
text(5, labels = "abline( h = h.CV.loglik$h.cv )", col = "gray60")
title("Bandwidth cross - validation plot")

```

In this plot it can be seen that the optimal bandwidth is around $2.5$, which is where the -loglikelihood is at its minimum. 


Now we make a local poison regression where le.fm.r is plotted against life.expectensy. 
```{r}
aux <- sm.poisson(countries$Life.expec,le.fm.r,h=h.CV.loglik$h.cv)
title(main=paste("h.cv.loglik = ",round(h.CV.loglik$h.cv,2),sep=""))

```





