---
title: "Local_poison_regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Bandwidth choice for the local Poisson regression

Modify the functions `h.cv.sm.binomial` and `loglik.CV` to obtain a bandwidth choice method for the local Poisson regression based on the loo-CV estimation of the expected likelihood of an independent observation. 

Remember that the leave-one-out cross-validation estimation of the expected log-likelihood of an independent observation, when using $h$ as bandwidth, is
\[
\ell_{CV}(h)
=\frac{1}{n}\sum_{i=1}^{n} \log\left( \widehat{\Pr}_h^{\tiny(-i)}(Y=y_i|X=x_i)\right),
\]
where $\widehat{\Pr}_h^{\tiny(-i)}(Y=y_i|X=x_i)$ is an estimation of
\[
\Pr(Y=y_i|X=x_i)=e^{-\lambda_i} \frac{\lambda_i^{y_i}}{y_i!},
\]
and 
\[
\lambda_i=\mathbb{E}(Y|X=x_i)
\]
should be estimated by maximum local likelihood using $h$ as bandwidth (for instance, using the function `sm.poisson` from the R package `sm`).
```{r}
# Bandwidth choice in the local logistic regression 
# by leave-one-out cross-validation.
# Function "sm.binomial", from library "sm", is used.
#
# Pedro Delicado

# method can be equal to 'loglik.CV' (default) or 'prob.missclas.CV'
h.cv.sm.poison <- function(x,y,rg.h=NULL,l.h=10,method=loglik.CV){
   cv.h <- numeric(l.h)
   if (is.null(rg.h)){
      hh <- c(h.select(x,y,method="cv"),
            h.select(x,y,method="aicc"))#,hcv(x,y))
      rg.h <- range(hh)*c(1/2, 2)
   }
   i <- 0
   gr.h <- exp( seq(log(rg.h[1]), log(rg.h[2]), l=l.h))
   for (h in gr.h){
      i <- i+1
      cv.h[i] <- method(x,y,h)
   }
   return(list(h = gr.h, 
               cv.h = cv.h, 
               h.cv = gr.h[which.min(cv.h)]))
}


# method loglik.CV: leave-one-out log-likelihood 
loglik.CV <- function(x,y,h){
  n <- length(x)
  pred <- sapply(1:n, 
      function(i,x,y,h){
         sm.poisson(x=x[-i],y=y[-i],h=h,eval.points=x[i],display="none")$estimate
      },   x,y,h)
  faculty <- sapply(1:n, function(i, y){prod(1:y[i])}, y)

  return(-sum(y*log(pred)-pred)) #-log(faculty) only a scale
}

```
```{r}
countries<-read.csv2(file="HDI.2017.subset.csv",row.names = 1)
#attach(countries)
le.fm.r <- round(countries$le.fm)
library(sm)

```

```{r}
x<- countries$Life.expec
y<- le.fm.r
h=500
i=3
n = length(x)
#sm.poisson(x=x[-i],y=y[-i],h=h,eval.points=x[i],display="none")$estimate
  pred <- sapply(1:n, 
      function(i,x,y,h){
         sm.poisson(x=x[-i],y=y[-i],h=h,eval.points=x[i],display="none")$estimate
      },   x,y,h)
sum(y*log(pred)-pred)


h.cv.sm.poison(Life.expec, le.fm.r)
```
```{r}
h.CV.loglik <- h.cv.sm.poison(Life.expec,le.fm.r,rg.h=c(1,3),method=loglik.CV)
plot(h.CV.loglik$h,h.CV.loglik$cv.h)
lines(h.CV.loglik$h,h.CV.loglik$cv.h)
```

```{r}
aux <- sm.poisson(Life.expec,le.fm.r,h=h.CV.loglik$h.cv)
title(main=paste("h.cv.loglik =",round(h.CV.loglik$h.cv,2),sep=""))

```












