---
title: "GAMs for Hirsutism data"
author: "Gregoire Gasparini, Aurora Hofman, Beatriu Tort"
date: "30 de marzo de 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hirsutism dataset

```{r}
hirs <- read.table("hirsutism.dat",header=T, sep="\t",fill=TRUE)
Treatment<- hirs$Treatment <- as.factor(hirs$Treatment)
summary(hirs)
```

#Boxplots to get an overview of the data. 
```{r}
boxplot(hirs[,2:5])
par(mfrow=c(2,2))
boxplot(hirs[,2]~hirs$Treatment,ylim=c(0,30), main=names(hirs)[2], xlab="Treatment")
boxplot(hirs[,3]~hirs$Treatment,ylim=c(0,30), main=names(hirs)[3], xlab="Treatment")
boxplot(hirs[,4]~hirs$Treatment,ylim=c(0,30), main=names(hirs)[4], xlab="Treatment")
boxplot(hirs[,5]~hirs$Treatment,ylim=c(0,30), main=names(hirs)[5], xlab="Treatment")
par(mfrow=c(1,1))
par(mfrow=c(2,2))
boxplot(hirs[hirs$Treatment==0,2:5],ylim=c(0,30), main="Treatment 0")
boxplot(hirs[hirs$Treatment==1,2:5],ylim=c(0,30), main="Treatment 1")
boxplot(hirs[hirs$Treatment==2,2:5],ylim=c(0,30), main="Treatment 2")
boxplot(hirs[hirs$Treatment==3,2:5],ylim=c(0,30), main="Treatment 3")
par(mfrow=c(1,1))

pairs(hirs)
```

###1st GAM model: linear model through GAM (FGm12 ~ Treatment + FGm0 + SysPres + DiaPres + weight + height)

```{r}
library(mgcv)

gam1 <- gam(FGm12 ~ Treatment + FGm0 + SysPres + DiaPres + weight + height, data = hirs)
summary(gam1)
```

###2nd GAM model: smooth model through GAM (FGm12 ~ Treatment + s(FGm0) + s(FGm0, by = Treatment) + s(SysPres) + s(DiaPres) + s(weight) + s(height))

We fit a full model with all the possible explanitory variables. 

```{r}
gam2 <- gam(FGm12 ~ Treatment + s(FGm0) + s(FGm0, by = Treatment) + s(SysPres) + s(DiaPres) + s(weight) + s(height), data = hirs)
summary(gam2)
```

From this model we can see that some variables potentially can be removed as they have a p-value that implies they are not significant.
Also the variable SysPres which is significant has a edf of 1 so it can be replaced by linear terms. 
```{r}
plot(gam2, residuals = TRUE, shade=TRUE, seWithMean=TRUE, pages = 7)
```

From this model we can see that some variables can be removed as they have a p-value that implies they are not significant.
Also the variable SysPres which is significant has a edf of 1 so it can be replaced by linear terms, this can also be seen from the plot. 
This leads us to the following model where s(DiaPres) + s(weight) + s(height) are removed and s(SysPres) is replaced by a linear term leading to a semiparametric model. 

###3rd GAM model: smooth model through GAM (FGm12 ~ Treatment + s(FGm0, by = Treatment) + SysPres)

```{r}
gam3 <- gam(FGm12 ~ Treatment + s(FGm0, by = Treatment) + SysPres, data = hirs) 
summary(gam3)
```

```{r}
plot(gam3, residuals = TRUE, shade=TRUE, seWithMean=TRUE)
```

#### Visualization of the joint effects of variables:

```{r}
vis.gam(gam3, view=c("Treatment","FGm0"), plot.type = "persp", theta=30, phi=30)
vis.gam(gam3, view=c("Treatment","FGm0"), plot.type = "contour")
```

```{r}
anova(gam3,gam2,test="F")
```

We see that the reduced model is rejected for the full model. 


###4rth GAM model: smooth model through GAM (FGm12 ~  s(FGm0, by = Treatment) + Treatment)


```{r}
gam4 <- gam(FGm12 ~ Treatment + s(FGm0, by = Treatment), data = na.omit(hirs)) 
summary(gam4)
anova(gam4, gam2, test = "F")
```

Again we reject the smaller model for the full model. 
Also we can see that almost all the s(FGm0):Treatment have a significant p value impying that we shoulf fit a model without these. 


```{r}
plot(gam4, residuals = TRUE, shade=TRUE, seWithMean=TRUE)
```

####Visualization of the joint effects of variables:

```{r}
vis.gam(gam4, view=c("Treatment","FGm0"), plot.type = "persp", theta=30, phi=30)
vis.gam(gam4, view=c("Treatment","FGm0"), plot.type = "contour")
```

###5th GAM model: smooth model through GAM (FGm12 ~ s(FGm0) + Treatment)

```{r}
gam5 <- gam(FGm12 ~ s(FGm0) + Treatment, data = na.omit(hirs))
summary(gam5)
anova(gam5, gam2, test = "F")
```
This model is also rejected compared to the full model.
```{r}
plot(gam5, residuals = TRUE, shade=TRUE, seWithMean=TRUE)
```

####Visualization of the joint effects of variables:

```{r}
vis.gam(gam5, view=c("Treatment","FGm0"), plot.type = "persp", theta=30, phi=30)
vis.gam(gam5, view=c("Treatment","FGm0"), plot.type = "contour")
```

###6th GAM model: smooth model through GAM (FGm12 ~ s(FGm0, by = Treatment) + SysPres + Treatment+ s(FGm0, SysPres)

```{r}
gam6 <- gam(FGm12 ~ s(FGm0, by = Treatment) + SysPres + Treatment+ s(FGm0, SysPres), data = hirs)
summary(gam6)
anova(gam6, gam2, test = "F")
```
Again the model we prefer is the first full model with all expenatory variables. 
```{r}
plot(gam6, residuals = TRUE, shade=TRUE, seWithMean=TRUE)
```

####Visualization of the joint effects of variables:

```{r}
vis.gam(gam6, view=c("SysPres","FGm0"), plot.type = "persp", theta=30, phi=30)
vis.gam(gam6, view=c("SysPres","FGm0"), plot.type = "contour")

vis.gam(gam6, view=c("Treatment","FGm0"), plot.type = "persp", theta=30, phi=30)
vis.gam(gam6, view=c("Treatment","FGm0"), plot.type = "contour")

vis.gam(gam6, view=c("Treatment","SysPres"), plot.type = "persp", theta=30, phi=30)
vis.gam(gam6, view=c("Treatment","SysPres"), plot.type = "contour")

```

## ANOVA type tests for the smaller models. 
```{r}
anova(gam1,gam2,test="F")
```

It is significant so model Gam2 explain better the variance than model 1. We conclude it is better to use a gam model than an ordinary linear model. 


```{r}
anova(gam4,gam3,test="F")
anova(gam5,gam4,test="F")
anova(gam5,gam6,test="F")
```

As we can see the the gam 4 is not rejected comparet to gam3. However gam5 is not rejectec compared to gam4 so we consider this a better model. Gam5 is also not rejected compared to gam6. 
This measn that the best model which is not the full model is gam5 which is FGm12 ~ s(FGm0) + Treatment. This is a very easy model and as we have seen earlier from the summary it only explains around 33 percent of the deviance. 

However it is seen like the full gam model which explains the most variability, also just explains around the 50% of the variablity and has a $R^2_{adj} = 0.368$.





