
---
title: "Local_poison_regression"
author: "Gregoire Gasparini, Aurora Hofman, Beatriu Tort"
date: "24 de marzo de 2020"
output: pdf_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(splines)
```

#Loading and reading data:
```{r}
load("bikes.Washington.Rdata")
```

```{r}
cnt <- bikes$cnt
instant <- bikes$instant
```

#Excersice 1 ----------------------------------------------------------

##Estimate function using smooth.spline:
```{r}
m<- smooth.spline(x= instant, y= cnt) #uses GCV by default. 
```

```{r}
#Penalty parameter:
m$lambda

#Correstponding degrees of freedom:
m$df

#Number of knots: 
length(m$fit$knot)
```

```{r}
df<- tibble(instant, cnt, m$x, m$y)

g<- ggplot(data = df, aes(instant, cnt)) + geom_point() + 
  geom_line(aes(m$x, m$y), colour = "red") + 
  ggtitle("Estimated function and original data")
g
```

##Estimation using R bs and lm
```{r}
n.knots <- m$df - 4
my.knots <- quantile(instant, ((1:n.knots)-.5)/n.knots)

m_bs <- bs(instant, knots = my.knots, intercept=T)
m_lm <- lm(cnt~m_bs-1)
```

```{r}
df_tot <- tibble(instant, cnt, m_lm$fitted.values, m$y)
g_lm <- ggplot(data = df_tot, aes(instant, cnt)) + 
  geom_point(colour = "red", shape= 1) + geom_line(aes(instant, m_lm$fitted.values), color = "blue") + 
  geom_line(aes(instant, m$y), colour = "green") + 
  ggtitle("Original data, cubic spline regression, unpenalized pline regrssion")
g_lm
```

Here is the color code :

- Red : initial data

- Green : smooth.spline() regression

- Blue : bs & lm regression


#Excersice 2 ---------------------------------------------------------------

##Use the function logistic.IRWLS.splines to fit the non-parametric binary regression cnt.5000 as a function of the temperature. 
```{r}
source('IRWLS_logistic_regression.R')

#Building the vector cnt.5000
cnt.5000 <- bikes$cnt > 5000
for (k in 1:length(cnt))
{
  if (cnt.5000[k] == TRUE) {cnt.5000[k] = 1} else {cnt.5000[k] = 0}
}

#First, We sort the data according to the explanatory variable.
x <- bikes$temp
y <- cnt.5000
sx <- sort(x,index.return =TRUE)
x <- sx$x
y <- y[sx$ix]

#Now we can fit the model
my.spline.glm <- 
  logistic.IRWLS.splines(x,y,
                         df=6, #df = 6 as demanded in instructions
                         all.knots=FALSE, plts = FALSE) 

#Plot
plot(x,y,col=2,xlab="temp",ylab="cnt.5000",main = "IRWLS logistic regression")
abline(v=my.knots,lty=2,col="grey")
lines(x,my.spline.glm$fitted.values,col=4)
abline(h=0.5)
```

##In which range of temperatures is $Pr(cnt>=5000|temp)$ larger than 0.5?
```{r}
probability = my.spline.glm$fitted.values

prob_above_50 <- (probability > 0.5)
indeces<-which(prob_above_50)
range<- c(x[indeces[1]], x[indeces[length(indeces)]])
range
```

The temperatures for which the chance of more than $5000$ bikes are rented is $20.74$ to $32.36$.

##Using the k-fold cross validation with k = 5 to choose df.
```{r}
source("locpolreg.R")

k.fold.cv <- function(x, y, k = 5, h = range(x)/10, p = 1, type.kernel = "normal"){
  n <- length(x)
  Ik <- floor((0:(n-1))/(n/k)) + 1
  ssr <- 0
  for (i in (1:k)){
    y.i <- y[Ik == i]
    aux <- locpolreg(x[Ik != i], y[Ik != i], h = h, p = p, tg = x[Ik == i],
                     type.kernel = type.kernel, doing.plot = FALSE)
    ssr <- ssr + sum((y.i-aux$mtgr)^2)
  }
  k.cv <- ssr/n
  return(k.cv)
}

h.k.fold.cv <- function(x, y, h.v = 3:15,
                        k = 5, p = 1, type.kernel = "normal"){
  n <- length(x)
  perm <- sample(1:n)
  xperm <- x[perm]
  yperm <- y[perm]
  
  k.cv <- h.v*0
  for (i in (1:length(h.v))){
    h <- h.v[i]
    k.cv[i] <- k.fold.cv(x = xperm, y = yperm, k = k, h = h, p = p,
                         type.kernel = type.kernel)
  }
  return(list(k = k, h.v = h.v, k.cv = k.cv))
}

bestdf <- h.k.fold.cv(x,y) 
h_opt<- bestdf$h.v[which.min(bestdf$k.cv)]
h_opt
```

The optimal df with a k-fold cross validation $k = 5$ is $3$.

##Recaulculing the non-parametric binary regression cnt.5000 as a function of temperature, using df = 3.
```{r}
#Now we can fit the model
my.spline.glm_h.df <- 
  logistic.IRWLS.splines(x,y,
                         df=3, #df = 3 as optimal df k-fold cross validation
                         all.knots=FALSE, plts = FALSE) 

#Plot
plot(x,y,col=2,xlab="temp",ylab="cnt.5000",main = "IRWLS logistic regression")
abline(v=my.knots,lty=2,col="grey")
lines(x,my.spline.glm_h.df$fitted.values,col=4)
abline(h=0.5)
```

##In this case the range of temperatures which $Pr(cnt>=5000|temp)$ is larger than 0.5.
```{r}
probability_h.df = my.spline.glm_h.df$fitted.values

prob_above_50_h.df <- (probability_h.df > 0.5)
indeces_h.df<-which(prob_above_50_h.df)
range_h.df<- c(x[indeces_h.df[1]], x[indeces_h.df[length(indeces_h.df)]])
range_h.df
```

With a $df = 3$ the range of the temperatures is $21.76$ to $35.33$

