---
title: "Estimating the conditional variance by local linear regression"
author: "Gregoire Gasparini, Aurora Hofman, Sarah Musiol, Beatriu Tort"
date: "17 de marzo de 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aircraft data
```{r}
library(sm)

data(aircraft)
help(aircraft)
lgPower <- log(aircraft$Power)
lgSpan <- log(aircraft$Span)
lgLength <- log(aircraft$Length)
lgWeight <- log(aircraft$Weight)
lgSpeed <- log(aircraft$Speed)
lgRange <- log(aircraft$Range)
Yr <- aircraft$Yr
```


## Estimating the conditional variance

### First, use the function loc.pol.reg that you can find in ATENEA and choose all the bandwidth values you need by leave-one-out cross-validation (you have not to program it again! Just look for the right function in the *.Rmd files you can find in ATENEA)

```{r}
source("locpolreg.R")
```

K Fold cross validation for leave-one-out cross validation from ATENEA:
```{r}
k.fold.cv <- function(x,y,k=10,h=range(x)/10,p=1,type.kernel="normal"){
  n <- length(x)
  Ik <- floor((0:(n-1))/(n/k))+1
  ssr <- 0
  for (i in (1:k)){
    y.i <- y[Ik==i]
    aux <- locpolreg(x[Ik!=i],y[Ik!=i],h=h,p=p,tg=x[Ik==i],
                     type.kernel=type.kernel, doing.plot=FALSE)
    ssr <- ssr + sum((y.i-aux$mtgr)^2)
  }
  k.cv <- ssr/n
  return(k.cv)
}

h.k.fold.cv <- function(x,y,h.v = exp(seq(log(diff(range(x))/20),
                                          log(diff(range(x))/4),l=10)), 
                        k=10,p=1,type.kernel="normal"){
  n <- length(x)
  perm <- sample(1:n)
  xperm <- x[perm]
  yperm <- y[perm]
  
  k.cv <- h.v*0
  for (i in (1:length(h.v))){
    h <- h.v[i]
    k.cv[i] <- k.fold.cv(x=xperm,y=yperm,k=k,h=h,p=p,
                         type.kernel=type.kernel)
  }
  return(list(k=k,h.v=h.v,k.cv=k.cv))
}

# In the pdf the teacher says to do a leave one out cross validation. This means to have a k=1, right? If it has to be one then the formula is not working. I am trying to figure it out.

bandwidth <- h.k.fold.cv(Yr, lgWeight)

bandwidth

```

```{r}
reg1 <- locpolreg(x=Yr, y=lgWeight, h=bandwidth$h.v) # I am not sure about h = ? I am not sure if it should be h.v or k.cv. I think should be k.cv but I am not sure.

reg1$mtgr

mx <- reg$mtgr

z <- log((lgWeight-mx)^2)

q <- locpolreg(Yr, z)

sigma <- exp(q$mtgr)

```

### Second, use the function sm.regression from library sm and choose all the bandwidth values you need by direct plug-in (use the function dpill from the same library KernSmooth).

```{r}
library(KernSmooth)
library(sm)

```

Just a test of the func with h arbritrary chosen
```{r}

#eval_points <- seq(6,13, length= 500)
sm.regression(x=lgWeight,y=Yr,h=.3)#, eval.points = eval_points)
#sm.regression(x=lgWeight,y=Yr,h=.3)
#sm.options(eval.grid =TRUE)
sm.regression(x=lgWeight,y=Yr,h=.3)

```

```{r}
h.dpill <- dpill(x=lgWeight,y=Yr,gridsize=101,
                 range.x=range(lgWeight))

print(h.dpill)
```
```{r}
fit_sm<- sm.regression(x=lgWeight,y=Yr,h=h.dpill, eval.points = eval_points)
```

```{r}
m_hat_sm <- fit_sm$estimate
```

Transform estimated residuals
```{r}
z <- log(fit_sm$se^2)

```

Just used the eval points form sm. We should se if we can chaneg them to beeing all the points of lgweights
```{r}

h.dpill_q <- dpill(x=fit_sm$eval.points,y=z,gridsize=101,
                 range.x=range(lgWeight))
q_hat_sm<- sm.regression(x=fit_sm$eval.points,y=z,h=h.dpill_q, eval.points=eval_points)
```
Estimate sigma^2
```{r}
sigma_squared_sm <- exp(q_hat_sm$estimate)
```

```{r}
bands <- 1.96*sqrt(sigma_squared_sm)
```

```{r}
library(ggplot2)

df <- data.frame(fit_sm$eval.points, fit_sm$estimate, bands, -bands)

g <- ggplot(df, aes(fit_sm$eval.points, fit_sm$estimate)) + geom_line() + geom_line(data = df, aes(fit_sm$eval.points, fit_sm$estimate+ bands), colour = 'red', linetype = "dashed")+ geom_line(data = df, aes(fit_sm$eval.points, fit_sm$estimate- bands), colour = 'red', linetype = "dashed")
g
```

