---
title: "Estimating the conditional variance by local linear regression"
author: "Gregoire Gasparini, Aurora Hofman, Sarah Musiol, Beatriu Tort"
date: "17 de marzo de 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aircraft data
```{r}
library(sm)

data(aircraft)
help(aircraft)
attach(aircraft)
lgPower <- log(Power)
lgSpan <- log(Span)
lgLength <- log(Length)
lgWeight <- log(Weight)
lgSpeed <- log(Speed)
lgRange <- log(Range)
```


## Estimating the conditional variance

### First, use the function loc.pol.reg that you can find in ATENEA and choose all the bandwidth values you need by leave-one-out cross-validation (you have not to program it again! Just look for the right function in the *.Rmd files you can find in ATENEA)

```{r}

```


### Second, use the function sm.regression from library sm and choose all the bandwidth values you need by direct plug-in (use the function dpill from the same library KernSmooth).

```{r}

```

