---
title: "Estimating the conditional variance by local linear regression"
author: "Gregoire Gasparini, Aurora Hofman, Sarah Musiol, Beatriu Tort"
date: "17 de marzo de 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aircraft data
```{r}
library(sm)

data(aircraft)
help(aircraft)
aircraft$lgPower <- log(aircraft$Power)
aircraft$lgSpan <- log(aircraft$Span)
aircraft$lgLength <- log(aircraft$Length)
aircraft$lgWeight <- log(aircraft$Weight)
aircraft$lgSpeed <- log(aircraft$Speed)
aircraft$lgRange <- log(aircraft$Range)
```


## Estimating the conditional variance

### First, use the function loc.pol.reg that you can find in ATENEA and choose all the bandwidth values you need by leave-one-out cross-validation (you have not to program it again! Just look for the right function in the *.Rmd files you can find in ATENEA)

```{r}
source("locpolreg.R")
```

K Fold cross validation for leave-one-out cross validation from ATENEA:
```{r}
k.fold.cv <- function(x,y,k=10,h=range(x)/10,p=1,type.kernel="normal"){
  n <- length(x)
  Ik <- floor((0:(n-1))/(n/k))+1
  ssr <- 0
  for (i in (1:k)){
    y.i <- y[Ik==i]
    aux <- locpolreg(x[Ik!=i],y[Ik!=i],h=h,p=p,tg=x[Ik==i],
                     type.kernel=type.kernel, doing.plot=FALSE)
    ssr <- ssr + sum((y.i-aux$mtgr)^2)
  }
  k.cv <- ssr/n
  return(k.cv)
}

h.k.fold.cv <- function(x,y,h.v = exp(seq(log(diff(range(x))/20),
                                          log(diff(range(x))/4),l=10)), 
                        k=10,p=1,type.kernel="normal"){
  n <- length(x)
  perm <- sample(1:n)
  xperm <- x[perm]
  yperm <- y[perm]
  
  k.cv <- h.v*0
  for (i in (1:length(h.v))){
    h <- h.v[i]
    k.cv[i] <- k.fold.cv(x=xperm,y=yperm,k=k,h=h,p=p,
                         type.kernel=type.kernel)
  }
  return(list(k=k,h.v=h.v,k.cv=k.cv))
}






```


### Second, use the function sm.regression from library sm and choose all the bandwidth values you need by direct plug-in (use the function dpill from the same library KernSmooth).

```{r}
library(KernSmooth)
```

